### 2023-10-27 chapters/background.tex
Updates

% A variational autoencoder (VAE) \parencite{kingma_autoencoding_2014,rezende_stochastic_2014} is a type of deep generative model and an instance of a latent variable model. 
In the following, we will introduce VAEs starting from the definition of latent variable models. We then derive the marginal likelihood objective used to train VAEs, discuss its different components, and highlight some central challenges of applying VAEs.

@@

\paragraph{Evidence lower bound}
By estimating the posterior distribution $p(\zb|\xb)$ with a variational approximation $q(\zb)$ we can derive a lower bound on the otherwise intractable marginal log-likelihood $\log p_\theta(\xb)$.
%
 \begin{align} \label{eq:variational-autoencoder-elbo}
    \log p_\thetab(\xb)
    &= \log \int p_\thetab(\xb,\zb),\textrm{d}\zb \nonumber \\
    &= \log \int q(\zb) \frac{p_\thetab(\xb,\zb)}{q(\zb)}\,\textrm{d}\zb \nonumber \\
    &\geq \int q(\zb) \log \left( \frac{p_\thetab(\xb,\zb)}{q(\zb)} \right) \,\textrm{d}\zb \nonumber \\
    &\geq \mathbb{E}_{q(\zb)} \left[ \log \frac{p_\thetab(\xb,\zb)}{q(\zb)} \right] \equiv \mathcal{L}_{}(\xb;q) \enspace .
    % &= \log \int p_\thetab(\xb|\zb)p(\zb),\textrm{d}\zb \nonumber \\
    % &= \log \int q(\zb) \frac{p_\thetab(\xb|\zb)p(\zb)}{q(\zb)}\,\textrm{d}\zb \nonumber \\
    % &\geq \int q(\zb) \log \left( \frac{p_\thetab(\xb|\zb)p(\zb)}{q(\zb)} \right) \,\textrm{d}\zb \nonumber \\
    % &\geq \mathbb{E}_{q(\zb)} \left[ \log \frac{p_\thetab(\xb|\zb)p(\zb)}{q(\zb)} \right] \equiv \mathcal{L}_{}(\xb;\thetab,\phib) \enspace .
\end{align}
%
Since we can evaluate the generative model $p_\thetab(\xb,\zb)$ and are free to define $q(\zb)$ as we please, this bound is tractable. 
However, finding an approximate posterior distribution $q(\zb)$ that is ...

\moretodo[inline]{Glue this together..}

Instead, VAEs amortize the cost of variational inference by parameterizing $q_\phib(\zb|\xb)$ as a single neural network. This makes it possible to efficiently compute approximate posterior distributions for any input $\xb$. The approximate posterior can also be used to compute a lower bound on the marginal distribution $p_\thetab(\xb)$ via Jensen's inequality, the so-called evidence lower bound (ELBO), 
%
\begin{equation}
    \log p_\thetab(\xb) \geq \mathbb{E}_{q_\phib(\zb|\xb)} \left[ \log \frac{p_\thetab(\xb,\zb)}{q_\phib(\zb|\xb)} \right] \equiv \mathcal{L}_{}(\xb;\thetab,\phib) \enspace .
\end{equation}
%
Although analytically evaluating the ELBO is still intractable, it can be approximated by sampling and used to train VAEs. We will return to this after examining the ELBO in more detail.

>>>>

% A variational autoencoder (VAE) \parencite{kingma_autoencoding_2014,rezende_stochastic_2014} is a type of deep generative model and an instance of a latent variable model. 
In the following, we will introduce VAEs starting from the definition of latent variable models. We then derive the marginal likelihood objective bound used to train them, discuss its different components and the approximation error, and highlight some central challenges of applying VAEs.

@@

\paragraph{Evidence lower bound}
By estimating the posterior distribution $p(\zb|\xb)$ with a variational approximation $q(\zb)$ we can derive a lower bound on the otherwise intractable marginal log-likelihood $\log p_\theta(\xb)$.
%
 \begin{align} \label{eq:variational-autoencoder-elbo}
    \log p_\thetab(\xb)
    &= \log \int p_\thetab(\xb,\zb),\textrm{d}\zb \nonumber \\
    &= \log \int q(\zb) \frac{p_\thetab(\xb,\zb)}{q(\zb)}\,\textrm{d}\zb \nonumber \\
    &\geq \int q(\zb) \log \left( \frac{p_\thetab(\xb,\zb)}{q(\zb)} \right) \,\textrm{d}\zb \nonumber \\
    &\geq \mathbb{E}_{q(\zb)} \left[ \log \frac{p_\thetab(\xb,\zb)}{q(\zb)} \right] \equiv \mathcal{L}_{}(\xb;q) \enspace .
    % &= \log \int p_\thetab(\xb|\zb)p(\zb),\textrm{d}\zb \nonumber \\
    % &= \log \int q(\zb) \frac{p_\thetab(\xb|\zb)p(\zb)}{q(\zb)}\,\textrm{d}\zb \nonumber \\
    % &\geq \int q(\zb) \log \left( \frac{p_\thetab(\xb|\zb)p(\zb)}{q(\zb)} \right) \,\textrm{d}\zb \nonumber \\
    % &\geq \mathbb{E}_{q(\zb)} \left[ \log \frac{p_\thetab(\xb|\zb)p(\zb)}{q(\zb)} \right] \equiv \mathcal{L}_{}(\xb;\thetab,\phib) \enspace .
\end{align}
%
Since we can evaluate the generative model $p_\thetab(\xb,\zb)$ and are free to define $q(\zb)$ as we please, this bound can be made tractable.

The usual approach in variational inference is to fit an approximate posterior $q(\zb)$ per data point. However, for large datasets, as is common in natural language processing, this can be too computationally expensive. 
% \moretodo[inline]{Glue this together..}
Instead, VAEs amortize the cost of variational inference by parameterizing the approximate posterior with a neural network that transforms any data point $\xb$ into the parameters of the corresponding conditional distribution, $q_\phib(\zb|\xb)$. This makes it possible to efficiently infer the approximate posterior distribution for any input $\xb$. 
With amortized variational inference, the ELBO can simply be written as,
% The approximate posterior can also be used to compute a lower bound on the marginal distribution $p_\thetab(\xb)$ via Jensen's inequality, the so-called evidence lower bound (ELBO), 
%
\begin{equation}
    \log p_\thetab(\xb) \geq \mathbb{E}_{q_\phib(\zb|\xb)} \left[ \log \frac{p_\thetab(\xb,\zb)}{q_\phib(\zb|\xb)} \right] \equiv \mathcal{L}_{}(\xb;\thetab,\phib) \enspace .
\end{equation}
%
Although analytically evaluating the ELBO is intractable due to the expectation, it can be approximated by sampling and then used to train VAEs. We will return to this after examining the ELBO in more detail.

### 2023-10-27 chapters/introduction.tex
Updates

One of the key advances needed to ensure that high-risk applications of machine learning comply with such requirements is accurate uncertainty estimation of their predictions. 
The multitude of rapid misclassifications made by the Uber self-driving car in Tempe is a likely indication that the model responsible for object-classification was overconfident in its predictions and unable to represent or communicate uncertainty to other systems in the car, or the safety driver. 
As a consequence, the model's many incoherent predictions were taken at face value and subsystems in the car acted on them without appropriate consideration to their reliability.
While any model will inevitably make wrong predictions, accurately estimating the associated uncertainty and communicating it to other systems or humans is necessary to reduce the risk that wrong predictions lead to catastrophic failures. 
Fittingly, a recent study by the European Parliamentary Research Services \parencite{europeanparliament_artificial_2022} concluded: 
%
\begin{quote}
    \centering\itshape
    Future AI solutions for healthcare should be implemented by integrating uncertainty estimation, a relatively new field of research that aims to provide clinicians with clinically useful indications on the degree of confidence in AI predictions.
\end{quote}
% \vspace{0.5em}
\begin{center}
\noindent\rule{0.2\textwidth}{0.5pt}
\end{center}
\vspace{1em}
%
% OUTLINE: Primer overview of the thesis. What are the topics and what is the context of the research.
% \noindent The research presented in this thesis will focus on the use of machine learning in relation to uncertainty estimation, speech processing, and medical conversations. 
\noindent The research presented in this thesis deals with the use of machine learning for medical interviews with a focus on uncertainty estimation and speech processing. 
The research project was partially funded by a grant from the national Danish Innovation Fund (grant no.\@ 0153-00167B) and defined and carried out in collaboration with the Danish company Corti, which develops computer software for medical decision support. 
In the rest of this chapter, we will provide a high-level introduction to these topics and to Corti and discuss the motivation behind the research project.

@@

\todo[inline]{Update here with:
I would try to embed a perspective for contemplation. Instead of us always talking about what the technology needs to improve on. Perhaps, it would also be prudent to ask an open-ended philosophical question: “While our requirements to the technology are in its place, one could ask, whether human beings have the ability to quantify uncertainty. In 2023, the prevalence of human errors in driving are much higher than self-driving vehicles.” <- Quick example. I’m sure you can find something better.}
% - a copilot for medical interviews
% - decision support
% - quality assurance
% - documentation
% -> many applications

>>>>

One of the key advances needed to ensure that high-risk applications of machine learning comply with such requirements is accurate uncertainty estimation of their predictions. 
The multitude of rapid misclassifications made by the Uber self-driving car in Tempe is a likely indication that the model responsible for object-classification was overconfident in its predictions and unable to represent or communicate uncertainty to other systems in the car, or the safety driver. 
As a consequence, the model's many incoherent predictions were taken at face value and subsystems in the car acted on them without appropriate consideration to their reliability.
While any model will inevitably make wrong predictions, accurately estimating the associated uncertainty and communicating it to other systems or humans is necessary to reduce the risk that wrong predictions lead to catastrophic failures. 
Fittingly, a recent study by the European Parliamentary Research Services \parencite{europeanparliament_artificial_2022} concluded: 
%
\begin{quote}
    \centering\itshape
    Future AI solutions for healthcare should be implemented by integrating uncertainty estimation, a relatively new field of research that aims to provide clinicians with clinically useful indications on the degree of confidence in AI predictions.
\end{quote}
% \vspace{0.5em}
\begin{center}
\noindent\rule{0.2\textwidth}{0.5pt}
\end{center}
\vspace{1em}
%
%
% OUTLINE: Primer overview of the thesis. What are the topics and what is the context of the research.
% \noindent The research presented in this thesis will focus on the use of machine learning in relation to uncertainty estimation, speech processing, and medical conversations. 
\noindent The research presented in this thesis deals with the use of machine learning for medical interviews with a focus on uncertainty estimation and speech processing. 
The research project was partially funded by a grant from the national Danish Innovation Fund (grant no.\@ 0153-00167B) and defined and carried out in collaboration with the Danish company Corti, which develops computer software for medical decision support. 
In the rest of this chapter, we will provide a high-level introduction to these topics and to Corti and discuss the motivation behind the research project.

@@

\moretodo[inline]{Update description of Corti. Focus on the current year. }

% - a copilot for medical interviews
% - decision support
% - quality assurance
% - documentation
% -> many applications
